Ecosystem size and complexity as extrinsic drivers of food chain length in branching ecosystems

### 2022-11-09 rmd/manuscript.Rmd
edit

---
title: "Ecosystem size and complexity are extrinsic drivers of food chain length in branching networks"
output:
  pdf_document:
    latex_engine: xelatex
  #word_document:
  #  reference_docx: word_format.docx
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
bibliography: references.bib
csl: ecosystems.csl
mainfont: Times New Roman
header-includes:
  \usepackage{textgreek}
---

@@

While recent research efforts of food web ecologists have substantially advanced our understanding of food chain drivers, most of the previous work has been conducted in simple, two-dimensional systems, such as lakes [@postEcosystemSizeDetermines2000; @doiResourceAvailabilityEcosystem2009] and oceanic islands [@takimotoEcosystemSizeNot2008; @pringlePredatorinducedCollapseNiche2019]. However, many natural habitats have a high degree of spatial complexity that cannot be represented simply by ecosystem size. Branching ecosystems, for example, are ubiquitous yet overlooked landscape structures (e.g., rivers, trees, caves, mountain ranges) [@grantLivingBranchesPopulation2007; @altermattDiversityRiverineMetacommunities2013]. While we know that food web interactions are spatially influenced [@calcagnoConstraintsFoodChain2011; @pillaiMetacommunityTheoryExplains2011; @takimotoEffectsProductivityDisturbance2012; @teruiSpatialDisturbanceSynchrony2019], the theory to account for it has lagged, particularly in branching habitat architectures. Here, we theoretically explore drivers of FCL in rivers, an excellent example of branching ecosystems.

@@

We generated branching networks as connected habitat patches (**Figure 1A**). The geometric arrangement of networks was controlled by two parameters: ecosystem size, the number of patches ($N_{p}$); and ecosystem complexity, branching probability ($P_{b}$). Habitat patches were assigned to be a confluence (or upstream terminal patch) with probability $P_{b}$, or an in-branch patch with probability $1-P_{b}$. Branches (tributaries) are a series of connected patches between confluences (or upstream terminal patches). The number of patches within branch $q$, $n_{p,q}$, is a random variable drawn from a geometric distribution $n_{p,q} \sim \text{Ge}(P_{b})$ but conditional on $N_{p} = \sum_{q}^{N_{b}} n_{p,q}$ ($N_{b}$ is the number of branches in a network). Full details of the network generation process is described in @teruiEmergentDualScaling2021.

In our simulation, disturbance occurs simultaneously with probability $p_m$ but habitat patches vary in their vulnerability. We simulated the spatial distribution of disturbance intensity (i.e., proportional mortality in the event of stochastic regional disturbance) as follows. A disturbance value, $m_{h}$, was assigned randomly to upstream terminal patch $h$ as $\text{logit}~m_{h} \sim \text{Normal}(\text{logit}~\mu_m, \sigma_{h}^2)$, where $\mu_m$ is the average of disturbance intensity at headwaters in an ordinary scale. The disturbance values propagate downstream through a spatial autoregressive process as:

$$
\text{logit}~m_{down} = \text{logit}~m_{up}+ \varepsilon
$$

@@

$$
\text{logit}~m_{down} = \omega~(\text{logit}~m_{1,up} +\varepsilon_1)+(1-\omega)~(\text{logit}~m_{2,up} +\varepsilon_2)
$$

@@

$$
K_{x} = K_{base}Q_{x}^z
$$

@@

We employed a discrete-time model to describe tri-trophic dynamics in each habitat patch. A discrete model is suitable for organisms with predictable reproduction cycles, a common life history for many lotic species, including macroinvertebrates and fish. Our simulation model consists of three species; a basal species $B$ with positive logistic growth up to a carrying capacity, $K$; a primary consumer species or intraguild prey, $C$, which grows based on the consumption of $B$; and an intraguild predator $P$, which can consume both $B$ and $C$ (**Figure 1C**). In our model, the realized species abundance $A_{i,x}(t+1)$ for species $i$ (either $B,C,~or~P$) at patch $x$ and time $t+1$ is expressed as a composite of within-patch trophic interactions, immigration, and emigration:

$$
A_{i,x}(t+1) \sim \text{Poisson}(\bar{A}_{i,x}(t) + I_{i,x}(t) - E_{i,x}(t))
$$

@@

***Internal dynamics.*** Each time step starts with internal dynamics. We accounted for background mortality, predator-induced mortality, and disturbance-induced mortality before the density-dependent growth occurs:

$$
\begin{aligned}
A'_{i,x} &= s_{0,i}~\beta_{i,x}(t)~A_{i,x}(t)\\
\bar{A}_{i,x}(t) &= g_i(A)~A'_{i,x}~\gamma_{x}(t)
\end{aligned}
$$

where $s_{0,i}$ is the background survival probability (density independent), $\beta_{i,x}(t)$ the survival probability after predation, and $\gamma_{x}(t)$ the survival probability after stochastic disturbance. The function $g_i(\centerdot)$ defines the form of the density-dependent recruitment for species $i$. Below, we explain how we described predation and recruitment processes for each species.

For the basal species, we considered the predation impacts of consumer and predator species:

$$
\beta_{B,x}(t) = 1 - [z_{BC,x}(t) + z_{BP,x}(t)]
$$
where $z_{BC,x}(t)$ and $z_{BP,x}(t)$ are the proportional mortality of basal species caused by consumer and predator species, respectively. The total mortality is calculated as the proportion of total individuals consumed ($W_{B,x}(t)$) relative to the current abundance of the basal species ($A_{B,x}(t)$):

$$
z_{BC,x}(t) + z_{BP,x}(t) =
\left\{
\begin{aligned}
&\frac{W_{B,x}(t)}{A_{B,x}(t)} &&\text{if $W_{B,x}(t) < A_{B,x}(t)$}\\
&1 &&\text{if $W_{B,x}(t) > A_{B,x}(t)$}\\
\end{aligned}
\right.
$$

@@

$$
W_{B,x}(t)=W_{BC,x}(t) + \delta_{x}(t) W_{BP,x}(t)
$$\
$W_{ij,x}(t)$ is the population-wide consumption of consumer $j$ on prey $i$ (number of individuals), and the parameter $\delta_{x}(t)$ is the relative preference of predator to basal species over consumer. We assume the Holling Type-II functional response for predator-prey interactions:

$$
W_{ij,x}(t) = \frac{a_{ij}A_{i,x}(t)A_{j,x}(t)}{A_{j,x}(t) + a_{ij}h_{ij}A_{i,x}(t)}
$$

@@

$$
\delta_{x}(t) = \frac{e_{BP} f_{BP}(A_{B,x}(t))}{e_{BP} f_{BP}(A_{B,x}(t)) + e_{CP} f_{CP}(A_{C,x}(t))}
$$

@@

$$
g_B(A) = \frac{r_{b}} {1 + \frac{r_b - 1}{K_{x}}A_{B,x}'}
$$

@@

For the consumer species, the predation-induced mortality is the proportion of individuals consumed by the predator species:

$$
\beta_{CP,x}(t) = 1 - z_{CP,x} = 1 - \frac{W_{CP,x}}{A_{C,x}(t)}
$$

@@

$$
g_C(A) =
\left\{
\begin{aligned}
&e_{BC}W_{BC,x}(t) &&\text{if $W_{B,x}(t) < A_{B,x}(t)$}\\
&e_{BC} \left( \frac{W_{BC,x}(t)}{W_{B,x}(t)} A_{B,x}(t) \right) &&\text{if $W_{B,x}(t) > A_{B,x}(t)$}\\\end{aligned}\right.
$$

@@

The predator species is not exposed to predation, thus $\beta_{P,x}(t) = 1$. The population growth of the predator considers two prey species:

$$
g_P(A) =
\left\{
\begin{aligned}
&e_{BP} \delta_{x}(t) W_{BP,x}(t) + e_{CP} (1 - \delta_{x}(t)) W_{CP,x}(t) &&\text{if $W_{B,x}(t) < A_{B,x}(t)$}\\
&e_{BP} \delta_{x}(t)\left(\frac{W_{BP,x}(t)}{W_{B,x}(t)}A_{B,x}\right) + e_{CP} (1 - \delta_{x}(t)) W_{CP,x}(t) &&\text{if $W_{B,x}(t) > A_{B,x}(t)$}\\
\end{aligned}
\right.
$$

After recruitment, the constituent species experience stochastic regional disturbance, which causes proportional loss of $m$ to populations:

$$
\gamma_{i,x}(t) = [1 - \psi(t) m_{x}]
$$

The binary random variable $\psi(t)$ expresses the occurrence of regional disturbance (disturbance occurs if $\psi(t)=1$ otherwise $0$), and $m_x$ is the disturbance intensity at patch $x$ as determined by the network generation process. $\psi(t)$ is drawn from a Bernoulli distribution with disturbance probability $p_m$.

@@

$$
\xi_{i,x}(t)=\frac{\sum_{y, y \ne x}E_{i,y}(t)\exp(-\theta d_{xy})}{\sum_{x} \sum_{y, y \ne x}\text{E}_{i,y}(t)\exp(-\theta d_{xy})}
$$

@@

$$
TP_{P,x}(t) = \delta_{x}(t) + 2(1 - \delta_{x}(t)) + 1
$$

@@

Before the main simulation, we performed an extensive sensitivity analysis to identify influential parameters (see **Supplementary Text** and **Table S1-S3**). Our sensitivity analysis revealed that ecosystem size and/or complexity effects were sensitive to disturbance, productivity, dispersal, and omnivory. Therefore, we ran the main simulation with `r n_scenario` parameter sets differing in these ecological contexts (two levels of disturbance, two levels of productivity, two levels of dispersal distance, and three levels of omnivory) and parameter values were summarized in **Table S1**.

@@

Under each scenario, we ran `r max(sim_main_result$n_rep)` simulation replicates with `r n_sim` timesteps of trophic dynamics (including `r n_sim - unique(sim_main_result$n_timestep)` steps of warmup and burn-in). During the warmup period (`r unique(sim_main_result$n_warmup)` time steps), we introduced propagules into habitat patches every 10 time steps to allow species to establish populations. Propagules are random draws from $Poisson(0.5)$. Subsequently, we ran a burn-in period with no propagules for `r unique(sim_main_result$n_burnin)` time steps to remove the effects of initial conditions. We obtained the mean FCL across all habitat patches for the last `r unique(sim_main_result$n_timestep)` time steps. Note that simulation replicates are differ in ecosystem size ($N_{p}\sim Unif(10,150)$) and ecosystem complexity ($P_{b} \sim Unif(0.01, 0.99)$) so we can estimate their effects on FCL. All analysis was performed in R 4.1.0 [@rcoreteamLanguageEnvironmentStatistical2021]

@@

Our simulation highlighted the interactive nature of putative controls of FCL. In particular, disturbance regimes strongly influenced the qualitative relationships between FCL and ecosystem properties. Below, we report our results for low and high disturbance regimes separately.

@@

In ecosystems with a low disturbance regime, omnivory largely determined FCL (colored lines in **Figure 2**), with no-omnivory food chains ($P$ only consumes $C$) having the longest FCL, and strong omnivory ($P$ consumes very little $C$) having the shortest FCL. The primary mechanism for the decreased FCL was different in low and high productivity scenarios. With low productivity, strong omnivory caused shorter FCL because it intensified competition between consumer and predator species. We can see this effect in the decreased occupancies of consumer and predator species in a low productivity scenario (most left column in **Figure S1-S4**). However, when productivity was high, the occupancies of both species were \~1.0, indicating that the structural constraint of strong omnivory made FCL shorter (second column from left in **Figure S1-S4**). Productivity played a secondary role in increasing FCL, particularly when omnivory was strong (compare left and right panels in **Figure 2A and 2B**). Dispersal had no apparent effects on FCL (compare top and bottom panels in **Figure 2A and 2B**).

@@

Dispersal buffered the disturbance effects on FCL and the magnitude of buffering differed by ecosystem structure. The greater buffering effect of dispersal emerged in larger ecosystems (compare top and bottom panels in **Figure 3**) because greater dispersal capability is required to reach distant patches. In the meantime, simpler, less branched ecosystems benefited more from increased dispersal because habitat patches are more isolated in linearly arranged networks (**Figure 3**). Productivity increased FCL as observed in a low disturbance regime, but it did not modulate the qualitative relationships between FCL and ecosystem properties (compare left and right panels in **Figure 3**).

Different processes drove the positive effects of ecosystem size and complexity on FCL when strong omnivory exists in low-productivity systems. In this scenario, the occupancies of $B$ and $C$ decreased at larger ecosystem sizes, while $P$ increased its occupancy (third column from left in **Figure S1 and S2**). Therefore, $P$ compensated for the loss of $B$ and $C$ to maintain the positive relationship between FCL and ecosystem size. In contrast, the increased occupancy of $C$ maintained the positive relationship between FCL and ecosystem complexity by offsetting the loss of $P$.

>>>>

---
title: "Ecosystem size and complexity as extrinsic drivers of food chain length in branching ecosystems"
output:
  pdf_document:
    latex_engine: xelatex
  #word_document:
  #  reference_docx: word_format.docx
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
bibliography: references.bib
csl: https://www.zotero.org/styles/ecosphere
mainfont: Times New Roman
header-includes:
  \usepackage{textgreek}
---

@@

While recent research efforts of food web ecologists have substantially advanced our understanding of food chain drivers, most of the previous work has been conducted in simple, two-dimensional systems, such as lakes [@postEcosystemSizeDetermines2000; @doiResourceAvailabilityEcosystem2009] and oceanic islands [@takimotoEcosystemSizeNot2008]. However, many natural habitats have a high degree of spatial complexity that cannot be represented simply by ecosystem size. Branching ecosystems, for example, are ubiquitous yet overlooked landscape structures (e.g., rivers, trees, caves, mountain ranges) [@grantLivingBranchesPopulation2007; @altermattDiversityRiverineMetacommunities2013]. While we know that food web interactions are spatially influenced [@calcagnoConstraintsFoodChain2011; @pillaiMetacommunityTheoryExplains2011; @takimotoEffectsProductivityDisturbance2012; @teruiSpatialDisturbanceSynchrony2019], the theory to account for it has lagged, particularly in branching habitat architectures. Here, we theoretically explore drivers of FCL in rivers, an excellent example of branching ecosystems.

@@

We generated branching networks as connected habitat patches (**Figure 1A;** see **Table 1** for parameter notations). The geometric arrangement of networks was controlled by two parameters: ecosystem size, the number of patches ($N_{p}$); and ecosystem complexity, branching probability ($P_{b}$). Habitat patches were assigned to be a confluence (or upstream terminal patch) with probability $P_{b}$, or an in-branch patch with probability $1-P_{b}$. Branches (tributaries) are a series of connected patches between confluences (or upstream terminal patches). The number of patches within branch $q$, $n_{p,q}$, is a random variable drawn from a geometric distribution $n_{p,q} \sim \text{Ge}(P_{b})$ but conditional on $N_{p} = \sum_{q}^{N_{b}} n_{p,q}$ ($N_{b}$ is the number of branches in a network). Full details of the network generation process is described in @teruiEmergentDualScaling2021.

In our simulation, disturbance occurs simultaneously across habitat patches with probability $p_m$ but the patches vary in their vulnerability. We simulated the spatial distribution of disturbance intensity (i.e., proportional mortality in the event of stochastic regional disturbance) as follows. A disturbance value, $m_{h}$, was assigned randomly to upstream terminal patch $h$ as $\text{logit}~m_{h} \sim \text{Normal}(\text{logit}~\mu_m, \sigma_{h}^2)$, where $\mu_m$ is the average of disturbance intensity at headwaters in an ordinary scale. The disturbance values propagate downstream through a spatial autoregressive process as:

$$
\begin{aligned}
\text{logit}~m_{down} &= \text{logit}~m_{up}+ \varepsilon &&\text{(1)}
\end{aligned}
$$

@@

$$
\begin{aligned}
\text{logit}~m_{down} &= \omega~(\text{logit}~m_{1,up} +\varepsilon_1)+(1-\omega)~(\text{logit}~m_{2,up} +\varepsilon_2) &&\text{(2)}
\end{aligned}
$$

@@

$$
\begin{aligned}
K_{x} &= K_{base}Q_{x}^z &&\text{(3)}
\end{aligned}
$$

@@

We employed a discrete-time model to describe tri-trophic dynamics in each habitat patch (see **Table 1** for parameter notations). A discrete model is suitable for organisms with predictable reproduction cycles, a common life history for many lotic species, including macroinvertebrates and fish. Our simulation model consists of three species; a basal species $B$ with positive logistic growth up to a carrying capacity, $K$; a primary consumer species or intraguild prey, $C$, which grows based on the consumption of $B$; and an intraguild predator $P$, which can consume both $B$ and $C$ (**Figure 1C**). In our model, the realized species abundance $A_{i,x}(t+1)$ for species $i$ (either $B,C,~or~P$) at patch $x$ and time $t+1$ is expressed as a composite of within-patch trophic interactions, immigration, and emigration:

$$
\begin{aligned}
A_{i,x}(t+1) &\sim \text{Poisson}(\bar{A}_{i,x}(t) + I_{i,x}(t) - E_{i,x}(t)) &&\text{(4)}
\end{aligned}
$$

@@

***Internal dynamics.*** Each time step starts with internal dynamics. We modeled density-dependent growth after accounting for background and predator-induced mortality:

$$
\begin{aligned}
\begin{aligned}
A'_{i,x} &= s_{0,i} \beta_{i,x}(t) A_{i,x}(t)\\
\bar{A}_{i,x}(t) &= g_i(A) A'_{i,x}
\end{aligned}
&&\text{(5)}
\end{aligned}
$$

where $s_{0,i}$ is the background survival probability (density independent) and $\beta_{i,x}(t)$ the survival probability after predation. The function $g_i(\cdot)$ defines the form of the density-dependent recruitment for species $i$. Below, we explain how we described predation and recruitment processes for each species.

*Basal species* -- We considered the predation impacts of consumer and predator species:

$$
\begin{aligned}
\beta_{B,x}(t) &= 1 - [z_{BC,x}(t) + z_{BP,x}(t)] &&\text{(6)}
\end{aligned}
$$ where $z_{BC,x}(t)$ and $z_{BP,x}(t)$ are the proportional mortality of basal species caused by consumer and predator species, respectively. The total mortality is calculated as the proportion of total individuals consumed ($W_{B,x}(t)$) relative to the current abundance of the basal species ($A_{B,x}(t)$):

$$
\begin{aligned}
z_{BC,x}(t) + z_{BP,x}(t) =
\left\{
\begin{aligned}
&\frac{W_{B,x}(t)}{A_{B,x}(t)} &&\text{if $W_{B,x}(t) < A_{B,x}(t)$}\\
&1 &&\text{if $W_{B,x}(t) > A_{B,x}(t)$}\\
\end{aligned}
\right.
&&\text{(7)}
\end{aligned}
$$

@@

$$
\begin{aligned}
W_{B,x}(t) &= W_{BC,x}(t) + \delta_{x}(t) W_{BP,x}(t) &&\text{(8)}
\end{aligned}
$$\
$W_{ij,x}(t)$ is the population-wide consumption of consumer $j$ on prey $i$ (number of individuals), and the parameter $\delta_{x}(t)$ is the relative preference of predator to basal species over consumer. We assume the Holling Type-II functional response for predator-prey interactions:

$$
\begin{aligned}
W_{ij,x}(t) &= \frac{a_{ij}A_{i,x}(t)A_{j,x}(t)}{A_{j,x}(t) + a_{ij}h_{ij}A_{i,x}(t)} &&\text{(9)}
\end{aligned}
$$

@@

$$
\begin{aligned}
\delta_{x}(t) = \frac{e_{BP} f_{BP}(A_{B,x}(t))}{e_{BP} f_{BP}(A_{B,x}(t)) + e_{CP} f_{CP}(A_{C,x}(t))}
&&\text{(10)}
\end{aligned}
$$

@@

$$
\begin{aligned}
g_B(A) = \frac{r_{b}} {1 + \frac{r_b - 1}{K_{x}}A_{B,x}'} &&\text{(11)}
\end{aligned}
$$

@@

*Consumer species* -- We modeled the predation-induced mortality as the proportion of individuals consumed by the predator species:

$$
\begin{aligned}
\beta_{CP,x}(t) = 1 - z_{CP,x} = 1 - \frac{W_{CP,x}}{A_{C,x}(t)} &&\text{(12)}
\end{aligned}
$$

@@

$$
\begin{aligned}
g_C(A) =
\left\{
\begin{aligned}
&e_{BC}W_{BC,x}(t) &&\text{if $W_{B,x}(t) < A_{B,x}(t)$}\\
&e_{BC} \left( \frac{W_{BC,x}(t)}{W_{B,x}(t)} A_{B,x}(t) \right) &&\text{if $W_{B,x}(t) > A_{B,x}(t)$}
\end{aligned}
\right.
&&\text{(13)}
\end{aligned}
$$

@@

*Predator species* -- The predator species is not exposed to predation, thus $\beta_{P,x}(t) = 1$. The population growth of the predator considers two prey species:

$$
\begin{aligned}
g_P(A) =
\left\{
\begin{aligned}
&e_{BP} \delta_{x}(t) W_{BP,x}(t) + e_{CP} (1 - \delta_{x}(t)) W_{CP,x}(t) &&\text{if $W_{B,x}(t) < A_{B,x}(t)$}\\
&e_{BP} \delta_{x}(t)\left(\frac{W_{BP,x}(t)}{W_{B,x}(t)}A_{B,x}\right) + e_{CP} (1 - \delta_{x}(t)) W_{CP,x}(t) &&\text{if $W_{B,x}(t) > A_{B,x}(t)$}\\
\end{aligned}
\right.
&&\text{(14)}
\end{aligned}
$$

***Disturbance.*** After recruitment, all species experience proportional loss of $m$ if stochastic regional disturbance occurs, modifying equation (5) as:

$$
\begin{aligned}
\bar{A}_{i,x}(t) &= g_i(A) A'_{i,x} (1 - m_{x}) &&\text{(15)}
\end{aligned}
$$

where $m_x$ is the disturbance intensity at patch $x$ as determined by the network generation process. Regional disturbance occurs with probability $p_m$.

@@

$$
\begin{aligned}
\xi_{i,x}(t) &= \frac{\sum_{y, y \ne x}E_{i,y}(t)\exp(-\theta d_{xy})}{\sum_{x} \sum_{y, y \ne x}\text{E}_{i,y}(t)\exp(-\theta d_{xy})}
&&\text{(16)}
\end{aligned}
$$

@@

$$
\begin{aligned}
TP_{P,x}(t) &= \delta_{x}(t) + 2(1 - \delta_{x}(t)) + 1
&&\text{(17)}
\end{aligned}
$$

@@

Before the main simulation, we performed an extensive sensitivity analysis to identify influential parameters (see **Supplementary Text** and **Table S1 -- S3**). Our sensitivity analysis revealed that ecosystem size and/or complexity effects were sensitive to disturbance, productivity, dispersal, and omnivory. Therefore, we ran the main simulation with `r n_scenario` parameter sets differing in these ecological contexts (two levels of disturbance, two levels of productivity, two levels of dispersal distance, and three levels of omnivory) and parameter values were summarized in **Table S1**.

@@

Under each scenario, we ran `r max(sim_main_result$n_rep)` simulation replicates with `r n_sim` timesteps of trophic dynamics (including `r n_sim - unique(sim_main_result$n_timestep)` steps of warmup and burn-in). During the warmup period (`r unique(sim_main_result$n_warmup)` time steps), we introduced propagules into habitat patches every 10 time steps to allow species to establish populations. Propagules are random draws from $\text{Poisson}(0.5)$. Subsequently, we ran a burn-in period with no propagules for `r unique(sim_main_result$n_burnin)` time steps to remove the effects of initial conditions. We obtained the mean FCL across all habitat patches for the last `r unique(sim_main_result$n_timestep)` time steps. Note that simulation replicates are differ in ecosystem size ($N_{p}\sim \text{Unif}(10,150)$) and ecosystem complexity ($P_{b} \sim \text{Unif}(0.01, 0.99)$) so we can estimate their effects on FCL. All analysis was performed in R 4.1.0 [@rcoreteamLanguageEnvironmentStatistical2021]

@@

## Overall pattern

There were three consistent patterns in the relationship between FCL and ecosystem properties. First, both ecosystem size and complexity had a stronger positive effect in high disturbance regime. Second, the effects of ecosystem complexity was stronger than ecosystem size in general. Third, omnivory had a negative influence on the FCL regardless of simulation scenarios. Below, we report our results for low and high disturbance regimes separately.

@@

In ecosystems with a low disturbance regime, omnivory largely determined FCL (colored lines in **Figure 2**), with no-omnivory food chains ($P$ only consumes $C$) having the longest FCL, and strong omnivory ($P$ consumes very little $C$) having the shortest FCL. The primary mechanism for the decreased FCL was different in low and high productivity scenarios. With low productivity, strong omnivory caused shorter FCL because it intensified competition between consumer and predator species. We can see this effect in the decreased occupancies of consumer and predator species in a low productivity scenario (most left column in **Figure S1 -- S4**). However, when productivity was high, the occupancies of both species were \~1.0, indicating that the structural constraint of strong omnivory made FCL shorter (second column from left in **Figure S1 -- S4**). Productivity played a secondary role in increasing FCL, particularly when omnivory was strong (compare left and right panels in **Figure 2A and 2B**). Dispersal had no apparent effects on FCL (compare top and bottom panels in **Figure 2A and 2B**).

@@

Dispersal buffered the disturbance effects on FCL and the magnitude of buffering differed by ecosystem structure. Larger ecosystems benefited more from long dispersal (compare top and bottom panels in **Figure 3**) because greater dispersal capability is required to reach distant patches. In particular, this pattern is clearer in simple, less branched ecosystems because habitat patches are more isolated in linearly arranged networks (**Figure 3**). Productivity increased FCL as observed in a low disturbance regime, but it did not modulate the qualitative relationships between FCL and ecosystem properties (compare left and right panels in **Figure 3**).

When strong omnivory exists in low-productivity systems, different processes drove the positive effects of ecosystem size and complexity on FCL. The positive effects of ecosystem size are mainly explained by the increased occupancy of $P$ (third column from left in **Figure S1 and S2**). In contrast, the increased occupancy of $C$ maintained the positive relationship between FCL and ecosystem complexity (third column from left in **Figure S3 and S4**).

### 2022-11-09 rmd/suppl_info.Rmd
edit

---
output:
  pdf_document:
    toc: true
    latex_engine: xelatex
    includes:
      before_body: title.tex
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
header-includes:
  \pagenumbering{gobble}
  \usepackage{amsmath}
---

@@

We performed a sensitivity analysis of the trophic community simulation to identify key simulation parameters that strongly affect the relationships between food chain length and ecosystem properties (the number of habitat patches $N_{p}$ and branching probability $P_{b}$. We generated 500 sets of parameter combinations by randomly drawing values of `r length(attributes(fit_sense$fit[[1]]$terms)$term.label)` simulation parameters from uniform distributions (**Table S1**). For each parameter combination, we generated `r max(sim_sv$n_rep)` branching networks with the gradients of ecosystem size ($N_p \sim Unif(10, 150)$) and complexity ($P_b \sim Unif(0.01, 0.99)$). This results in a total of `r nrow(sim_sv)` simulation replicates. We ran `r n_sim_stvy` time steps of trophic community dynamics in each simulation replicate. We obtained temporal means of food chain length for the last `r unique(sim_sv$n_timestep)` time steps. The first `r n_sim_stvy - unique(sim_sv$n_timestep)` time steps were discarded as initialization and burn-in periods.

@@

```{=tex}
\begin{align}
  u_{n} &\sim Normal(\mu_n, \sigma^2)\\
  \mu_n &= \zeta_0 + \sum_{k}{\zeta_k x_{k, n}}
\end{align}
```
where $\zeta_k$ are the intercept and regression coefficients, and $x_{k,n}$ are the simulation parameters drawn from uniform distributions (**Table S1**). Response and explanatory variables were standardized to a mean of zero and a standard deviation of one so that regression coefficients are comparable.

>>>>

---
output:
  pdf_document:
    toc: true
    latex_engine: xelatex
    includes:
      before_body: title.tex
mainfont: Times New Roman
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
header-includes:
  \pagenumbering{gobble}
  \usepackage{amsmath}
---

@@

We performed a sensitivity analysis of the trophic community simulation to identify key simulation parameters that strongly affect the relationships between food chain length and ecosystem properties (the number of habitat patches $N_{p}$ and branching probability $P_{b}$. We generated 500 sets of parameter combinations by randomly drawing values of `r length(attributes(fit_sense$fit[[1]]$terms)$term.label)` simulation parameters from uniform distributions (**Table S1**). For each parameter combination, we generated `r max(sim_sv$n_rep)` branching networks with the gradients of ecosystem size ($N_p \sim \text{Unif}(10, 150)$) and complexity ($P_b \sim \text{Unif}(0.01, 0.99)$). This results in a total of `r nrow(sim_sv)` simulation replicates. We ran `r n_sim_stvy` time steps of trophic community dynamics in each simulation replicate. We obtained temporal means of food chain length for the last `r unique(sim_sv$n_timestep)` time steps. The first `r n_sim_stvy - unique(sim_sv$n_timestep)` time steps were discarded as initialization and burn-in periods.

@@

```{=tex}
\begin{align}
  u_{n} &\sim \text{Normal}(\mu_n, \sigma^2)\\
  \mu_n &= \zeta_0 + \sum_{k}{\zeta_k x_{k, n}}
\end{align}
```
where $\zeta_k$ are the intercept and regression coefficients, and $x_{k,n}$ are the simulation parameters drawn from uniform distributions (**Table S1**). Response and explanatory variables were standardized to a mean of zero and a standard deviation of one so that regression coefficients are comparable.
